Model complexity and model choice for animal movement models

### 2016-06-08 guelph_talk/guelph_panther.Rnw
talk cont'd

\title[Animal movement]{Model complexity and model choice for animal movement models}
\author[Ben Bolker]{Ben~Bolker, McMaster University \\
  {\tiny Departments of Mathematics \& Statistics and Biology}}
\institute{Guelph Biomathematics \& Biostatistics Symposium}
\date{9 June 2016}
% \pgfdeclareimage[height=0.5cm]{uflogo}{letterhdwm}
% \logo{\pgfuseimage{uflogo}}
\AtBeginSubsection[]{
  \frame<beamer>{ 
     \frametitle{Outline}   
     \tableofcontents[currentsection,currentsubsection] 
   }
 }

@@

\begin{frame}
\frametitle{To do}
\begin{itemize}
\item pix!
\item general REFS: Turchin, Morales et al, Langrock ... ?
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item basic biology: males move faster, farther
\item three states are identifiable, sensible
\item ...
\end{itemize}

@@

\begin{frame}
\frametitle{Animal movement: open challenges}
\begin{itemize}
\item Cognition
\item Intraspecific interaction/collective movement
\item Continuous-time movement models
\item Edges, barriers, and corridors
\item Efficient (big-data) approaches
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item simple model families +  \\
  model misspecification $\to$ \\
  overparameterization

\end{itemize}
\end{frame}

@@

>>>>

\title[Animal movement]{Model complexity and model choice for animal movement models}
\author[Ben Bolker]{Ben~Bolker, McMaster University \\
  {\tiny Departments of Mathematics \& Statistics and Biology}}
\institute{Guelph Biomathematics \& Biostatistics Symposium}
\date{9 June 2016}
% \pgfdeclareimage[height=0.5cm]{uflogo}{letterhdwm}
% \logo{\pgfuseimage{uflogo}}
\AtBeginSection[]{
  \frame<beamer>{ 
     \frametitle{Outline}   
     \tableofcontents[currentsection,currentsubsection] 
   }
 }

@@

@@

\begin{itemize}
\item basic biology: males move faster, farther
\item three states are identifiable, sensible
\item dwell distributions approximately geometric \\
(HSMM $\to$ HMM)
\end{itemize}

@@

\begin{frame}
\frametitle{Animal movement: open challenges}
\begin{itemize}
\item Cognition
\item Intraspecific interaction/collective movement \citepx{delgado_statistical_2014}
\item Continuous-time movement models \citepx{calabrese_ctmm_2016}
\item Edges, barriers, and corridors \citepx{beyer_you_2016}
\item Efficient (big-data) approaches \citepx{brillinger_three_2008}
\item Putting it all together ...
\end{itemize}
\end{frame}

@@

\begin{itemize}
\item simple model families +  \\
  model misspecification $\to$ \\
  overparameterization
\item protocols and tools for model checking \citepx{potts_generalized_2014};
score tests?
\item flexible computational frameworks
\end{itemize}
\end{frame}

@@

\begin{columns}
\begin{column}{0.5\textwidth}
\begin{itemize}
\item ``should I use AIC or BIC? I heard that AIC is inconsistent ...''
\item complexity penalty = 2 (AIC) vs $\log(n)$ (BIC)
\item best prediction vs. model identification \citepx{yang_can_2005}
\item \emph{effect size spectrum}: \\
tapering or discrete?
\end{itemize}
\end{column}
\begin{column}{0.5\textwidth}
<<effectsize,echo=FALSE,fig.height=5,fig.width=4>>=
d <- data.frame(type=rep(c("tapering","discrete"),each=10),
                n=rep(1:10,2),
                val=c(0.6^(0:9),rep(c(1,0),c(3,7))))
ggplot(d,aes(n,val,fill=type))+
    geom_bar(stat="identity",position="identity",alpha=0.6)+
    theme(axis.ticks=element_blank(),
          axis.text.x=element_blank(),
          axis.text.y=element_blank())+
    labs(x="effect",y="effect size")+
    scale_fill_brewer(palette="Set2")
@ 
\end{column}
\end{columns}
\end{frame}
